---
title: "Tutorial for the `pridr` software package, prepared for IAMC 2022"
authors: Kanishka B. Narayan, Brian C. O'Neill, Stephanie Waldhoff, Claudia Tebaldi, Kelly C. Casper
  & Kelly C. Casper
output:
   html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    #toc_depth: 3
    #number_sections: True
    theme: cosmo
    #inverse: true

---
## Pre-processing, loading libraries

```{r setup, warning=FALSE, message=FALSE}
##LOAD our package
library(pridr)


#Other packages
library(ggplot2)
library(ggsci)
library(parallel)
library(tidyr)
library(dplyr)
library(data.table)
library(assertthat)
library(viridis)
library(knitr)


#Basic format for figs (optional)
scheme_basic <- theme_bw() +
  theme(legend.text = element_text(size = 15)) +
  theme(legend.title = element_text(size = 15)) +
  theme(axis.text = element_text(size = 18)) +
  theme(axis.title = element_text(size = 18, face = "bold")) +
  theme(plot.title = element_text(size = 15, face = "bold", vjust = 1)) +
  theme(plot.subtitle = element_text(size = 9, face = "bold", vjust = 1))+ 
  theme(strip.text = element_text(size = 7))+
  theme(strip.text.x = element_text(size = 18, face = "bold"))+
  theme(strip.text.y = element_text(size = 15, face = "bold"))+
  theme(legend.position = "bottom")+
  theme(legend.text = element_text(size = 12))+
  theme(legend.title = element_text(size = 12,color = "black",face="bold"))+
  theme(axis.text.x= element_text(hjust=1))+
  theme(legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"))
```

## Example 1: Generate deciles using lognormal approach

### `We load a sample dataset with GINI coefficients and mean income (by country, year) and derive the income distribution (deciles) using the lognormal based approach.`

#### `1. Load a sample dataset`
```{r, warning=FALSE, message=FALSE}

read.csv("Input_Data/Wider_aggregated_deciles.csv", stringsAsFactors = FALSE) %>%
  select(country, year, gdp_ppp_pc_usd2011, gini) %>%
  distinct() %>%
  mutate(sce="Historical data") %>% 
  filter(year > 2013)->data_for_lognorm

knitr::kable(head(data_for_lognorm), format = "html")
```

#### `2. Use the lognormal model on this dataset`
```{r, warning=FALSE, message=FALSE,echo=TRUE}
start_time= Sys.time()
compute_deciles_lognormal(data_for_lognorm)->lognormal_model

end_time=Sys.time()

print(paste0("Processed in ",as.integer(end_time-start_time), " seconds"))

knitr::kable(head(lognormal_model), format = "html")

```

## Example 2: Generate lognormal density dist (For an abritary set of parameters)

### `Here, we generate lognormal density distribution from arbritary values of the GINI and mean income.`

```{r, warning=FALSE, message=FALSE}

density_dist_1 <- compute_lognormal_dist(mean_income = 15,gini=0.6) %>% mutate(gini=as.character(0.6))
density_dist_2 <- compute_lognormal_dist(mean_income = 15,gini=0.3)%>% mutate(gini=as.character(0.3))
density_dist_3 <- compute_lognormal_dist(mean_income = 15,gini=0.2)%>% mutate(gini=as.character(0.2))

g <- ggplot(data=bind_rows(density_dist_3,density_dist_2,density_dist_1), aes(x=gdp_pcap,y=density,color=gini))+
     geom_line(size =2)+scale_color_aaas()+ xlab("Income in Thous USD")

g+scheme_basic

```

## Example 3. Use the PCA based model 

### `Here , we demonstrate the usage of the PCA based model. We pass a dataset of required variables (Step 1 below) to the `PC_model` function (Step 2 below). Note that the function parameters are stored in memory for pridr and will not change unless the user specifically resets parameters (using newer data for example).`

##### `1. Compile data (This is the dataset format that needs to be passed to the function)`
```{r, warning=FALSE, message=FALSE}
sample_data <- read.csv("Input_Data/sample_data.csv")
knitr::kable(head(sample_data), format = "html")

```
#### `2. Run PC model on this dataset`
```{r, warning=FALSE, message=FALSE}
start_time <- Sys.time()

PC_model_results <- PC_model(sample_data %>% filter(iso=="usa"))


print(paste0("Completed in ", as.integer(Sys.time()-start_time), " seconds."))

knitr::kable(head(PC_model_results), format = "html")


```
#### `3. Plot results`
```{r, warning=FALSE, message=FALSE}
g <- ggplot(data=PC_model_results %>% filter(year %in% c(2010:2015)) ,aes(x=factor(Category,levels =c('d1','d2','d3','d4',
  'd5','d6','d7','d8',                                                                                'd9','d10')),y=pred_shares,color=year,group=year))+
     geom_line()+
     geom_point()+scale_color_viridis()+xlab("Deciles")+
     ggtitle("Income distributions in the US in multiple years  ")

g+scheme_basic+theme(legend.position = "right")
    
```

## Example 4. Generate GINI coefficients for a given set of deciles

### ` `pridr` allows users to recalculate summary metrics such as the GINI coefficient from deciles.`

```{r, warning=FALSE, message=FALSE}
gini_data <-compute_gini_deciles(PC_model_results %>% mutate(category=Category), inc_col = "pred_shares",grouping_variables = c("country","year")) %>% rename(gini=output_name)
knitr::kable(head(gini_data), format = "html")
```

## Example 5. Aggregate deciles to a region (Here we aggregate country level distributions to a global distribution)

### `Users can also aggregate the income distribution to a regional or global scale based on arbritary mapping of ISOs to regions. Here, we aggregate all income distributions at the country level to a global income distribution in 2015.`


#### `1. Compile data`
```{r, warning=FALSE, message=FALSE}
read.csv("Input_Data/Wider_aggregated_deciles.csv", stringsAsFactors = FALSE) %>%
  select(country, year, gdp_ppp_pc_usd2011, population,Income..net.,Category) %>% filter(year%in% c(2015)) %>% mutate(GCAM_region_ID="Global")->ISO_data
knitr::kable(head(ISO_data), format = "html")

```

#### `2. Aggregate the distribution`
```{r, warning=FALSE, message=FALSE}


aggregate_country_deciles_to_regions(ISO_data,
                                     grouping_variables = c("GCAM_region_ID","year")
                                     )->agg_data

knitr::kable(head(agg_data), format = "html")
```

#### `3. Plot the distributions`

```{r, warning=FALSE, message=FALSE}
g <- ggplot(data=ISO_data,aes(x=factor(Category,levels =c('d1','d2','d3','d4',
  'd5','d6','d7','d8',                                                                                'd9','d10')),y=Income..net.),group=country)+
     geom_line(data=ISO_data,aes(x=factor(Category,levels =c('d1','d2','d3','d4',
  'd5','d6','d7','d8',                                                                                'd9','d10')),y=Income..net.,group=country),color="grey")+
     geom_point(color="grey")+scale_color_aaas()+xlab("Deciles")+facet_wrap(~year)+
     geom_line(data=agg_data, aes(x=factor(category,levels =c('d1','d2','d3','d4',
  'd5','d6','d7','d8',                                                                                'd9','d10')),y=shares),color="black",group=agg_data$GCAM_region_ID,size=2.5)+
     ylab("shares")+
     xlab("Deciles")+
  ggtitle("Grey lines are country level distributions, black line is \naggregated global distribution in 2015")

g+scheme_basic
```

